---
title: "Exploring CQI Coffee data"
author: "Simon Laursen & Hjalte With"
date: " created 09 January 2022, updated r Sys.Date()"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading libraries
```{r}
library(tidyverse)
library(lubridate)
library(janitor)
library(dplyr)
library(ggcorrplot) 
library(forcats)

```


# Loading data and transforming empty cells to NA
```{r}
cqi_null_too_na <-  read.csv("~/OneDrive/UNI/5. semester/CDS/CDS_project/merged_data_cleaned.csv",  na.strings=c("","NA"))
```


# Cleaning the data
```{r}
#removing unnecessary columns
cqi_remove_col <- cqi_null_too_na[-c(1,3,5:18,33,34,36:40)] 

#Cleaning variables regarding altitude. 
#Converting every variables from feet to meter, by detecting all values equal to ft in the variable unit of measurement. 
#Afterwards we convert feet to meter by multiplying the corresponding value in the variable altitude_mean_meters with 0.3048.
ft_to_m <- cqi_remove_col %>% 
  mutate(altitude_mean_meters = case_when(
    str_detect(unit_of_measurement, "ft") ~ altitude_mean_meters * 0.3048,
    TRUE ~ altitude_mean_meters)
  )

#Removing altitude values above 4000 m.
rem_high_alt <- ft_to_m %>% 
  subset(altitude_mean_meters < 4000)

#Removing remaining altitude variables except altitude_mean_meters
rem_alt_vars <- rem_high_alt[-c(18:20)]

#Creating final name for the clean data set
coffee_data <- rem_alt_vars





      
```



# Bar plots for data exploration
```{r}


#Distribution of coffee species
#Counting the occurrence of the two different coffee species values and plotting them in a bar plot with ggplot.
ggplot(coffee_data, aes(Species))+
    geom_bar(fill="#f68060", alpha=.6, width=.4) +
    coord_flip() +
    theme_bw()

#Distribution of Processing Method 
#Counting the occurrence of the different Processing methods and omiting all NA's
proces_method_count <- coffee_data %>% 
  na.omit(Processing.Method) %>% 
  count(Processing.Method)

#Plotting the occurrence of the different values in descending order
proces_method_count %>% 
  mutate(Processing.Method = fct_reorder(Processing.Method, n)) %>%
  ggplot(aes(Processing.Method, n))  +
    geom_bar(stat="identity", fill="#f68060", alpha=.6, width=.4) +
    coord_flip() +
    xlab("") +
    theme_bw()

#Distribution of Variety of coffee sort
#Counting the occurrence of the different varieties of coffee and omiting all NA's
variety_count <- coffee_data %>% 
  na.omit(Variety) %>% 
  count(Variety)

#Plotting the occurrence of the different values in descending order
variety_count %>% 
  mutate(Variety = fct_reorder(Variety, n)) %>%
  ggplot(aes(Variety, n))  +
    geom_bar(stat="identity", fill="#f68060", alpha=.6, width=.4) +
    coord_flip() +
    xlab("") +
    theme_bw()

#Distribution of Country of origin
#Counting the occurrence of different countries
country_count <- coffee_data %>% 
  count(Country.of.Origin)

#Plotting the occurrence of the different values in descending order
country_count %>% 
  mutate(Country.of.Origin = fct_reorder(Country.of.Origin, n)) %>%
  ggplot(aes(Country.of.Origin, n))  +
    geom_bar(stat="identity", fill="#f68060", alpha=.6, width=.4) +
    coord_flip() +
    xlab("") +
    theme_bw()
```
# Boxplots to examine coherence between diffent variables
```{r}
#Boxplot showing coherence between species and TCP

coffee_data %>% 
 ggplot(aes(x = Species, y = Total.Cup.Points)) +
    geom_boxplot() +
    #geom_jitter(color="black", size=0.4, alpha=0.9) +
    coord_flip() + 
    ylim(60,90)
    ggtitle("Species effect on Total Cup Points")
    
ggsave("plot3.png", height = 10, width = 20, dpi = 200)

#Boxplot showing coherence between countries and TCP

coffee_data %>% 
 ggplot(aes(x = Country.of.Origin, y = Total.Cup.Points)) +
    geom_boxplot() +
    geom_jitter(color="black", size=0.4, alpha=0.9) +
    coord_flip() +                                    
    ylim(60,100) +                                     
    ggtitle("Contry of Origin effect on Total Cup Points")


ggsave("plot2.png", height = 10, width = 10, dpi = 200)

```

# Distribution of country vs Mean TCP for each country
```{r}

#mean TCP for each country
mean_TCP <- coffee_data %>% 
  group_by(Country.of.Origin) %>%   #Grouping the dataset by Country of origin
  summarise_at(vars(Total.Cup.Points),funs(mean(.))) %>% #Summarizing at the TCP variable and using funs to get the average TCP-score for each country. 
  mutate_if(is.numeric, round, digits=2) %>% #Rounding to 2 decimals
  arrange(desc(Total.Cup.Points)) #Arranging the list in descending order
mean_TCP


#Arranging top 10 countries by occurrence (pulling variable name countrycount from previous section)
top10_country_count <- country_count %>% 
  arrange(desc(n))
head(top10_country_count, 10)


#Plotting top 10 countries by average TCP. We have chosen these country values based on top10_country_count 
  ggplot(subset(mean_TCP, Country.of.Origin %in% c("Mexico", "Colombia", "Guatemala", "Brazil", "Taiwan", "United States (Hawaii)", "Honduras", "Costa Rica", "Ethiopia", "Tanzania, United Republic Of")),
   aes(x = Country.of.Origin, y = Total.Cup.Points)) + 
   geom_bar( stat = "identity", width = 0.2, position = "dodge") + 
   geom_text(aes(label = Total.Cup.Points, y = Total.Cup.Points + 3),
             position = position_dodge(0.9),
             vjust = 0.25) +
   coord_flip() +
   ylim(0,90) +
   labs( title = "Mean TCP for top 10 countries based on amount of reviews", x = "Country of origin", y = "Mean total cup point")
  
ggsave("plot1.png", height = 5, width = 10, dpi = 200)


# Merging occurrence and average TCP of top 10 countries
merge_TCP_count <- merge(mean_TCP, top10_country_count, by = "Country.of.Origin")

#Plotting merged data frame and coloring by country of origin
ggplot(subset(merge_TCP_count, 
    Country.of.Origin %in% c("Mexico", "Colombia", "Guatemala", "Brazil", "Taiwan", "United States (Hawaii)", "Honduras", "Costa Rica", "Ethiopia", "Tanzania, United Republic Of")), 
    aes(x=Total.Cup.Points, y=n, color = Country.of.Origin))+
    geom_point()


```
# Distribution of TCP's level of altitude
```{r}
#Plotting TCP and altitude
ggplot(coffee_data, aes(x=altitude_mean_meters, y=Total.Cup.Points)) +
    geom_point(alpha=0.7) +
    ylim(55, 100) #Limiting the y-axis to start at 55

    ggsave("plot4.png", height = 15, width = 20, dpi = 200)
    
    
```

# Coherence between TCP and Average Altitude for each country

```{r}

# Plotting every observations TCP and altitude value and coloring by country
ggplot(subset(coffee_data, 
    Country.of.Origin %in% c("Mexico", "Colombia", "Guatemala", "Brazil", "Taiwan", "United States (Hawaii)", "Honduras", "Costa Rica", "Ethiopia", "Tanzania, United Republic Of")), 
    aes(x=Total.Cup.Points, y=altitude_mean_meters, color = Country.of.Origin)) +
    geom_point() +
    xlim(75, 92) +
    labs(x = "Total Cup Points", y = "Average Altitude (m)")
    
```


#Correlation of grading variables
```{r}


# Choosing specific grading variables to include in new data frame
grading_correlation <- coffee_data[c(5:15)]

# Computing the correlation between all variables in grading_correlation
corr <- cor(round((grading_correlation),1))
head(corr[1:10])

# Plotting the correlation data frame grading_correlation
ggcorrplot(corr,
  hc.order = TRUE,
  type = "lower",
  lab = TRUE,
  method = "square",
  outline.color = "white",
  ggtheme = ggplot2::theme_gray,
  colors = c("#6D9EC1", "white", "#4633FF"),
  ) + 
  theme(
    legend.key.height = unit(1.5,"cm")
  )

ggsave("plot5.png", height = 10, width = 20, dpi = 200)
 
```

